\TODO{What's the title?}

### 2018-09-18 doc/Paper.tex
Update PDF to 43d2717eec03c9db275147b4e66fb12fc705110f

\EndFmtInput
\makeatother
%
%
%
%
%
%
% This package provides two environments suitable to take the place
% of hscode, called "plainhscode" and "arrayhscode". 
%
% The plain environment surrounds each code block by vertical space,
% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
% similar to formulas. Note that if these dimensions are changed,
% the spacing around displayed math formulas changes as well.
% All code is indented using \leftskip.
%
% Changed 19.08.2004 to reflect changes in colorcode. Should work with
% CodeGroup.sty.
%
\ReadOnlyOnce{polycode.fmt}%
\makeatletter

\newcommand{\hsnewpar}[1]%
  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}

% can be used, for instance, to redefine the code size, by setting the
% command to \small or something alike
\newcommand{\hscodestyle}{}

% The command \sethscode can be used to switch the code formatting
% behaviour by mapping the hscode environment in the subst directive
% to a new LaTeX environment.

\newcommand{\sethscode}[1]%
  {\expandafter\let\expandafter\hscode\csname #1\endcsname
   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}

% "compatibility" mode restores the non-polycode.fmt layout.

\newenvironment{compathscode}%
  {\par\noindent
   \advance\leftskip\mathindent
   \hscodestyle
   \let\\=\@normalcr
   \let\hspre\(\let\hspost\)%
   \pboxed}%
  {\endpboxed\)%
   \par\noindent
   \ignorespacesafterend}

\newcommand{\compaths}{\sethscode{compathscode}}

% "plain" mode is the proposed default.
% It should now work with \centering.
% This required some changes. The old version
% is still available for reference as oldplainhscode.

\newenvironment{plainhscode}%
  {\hsnewpar\abovedisplayskip
   \advance\leftskip\mathindent
   \hscodestyle
   \let\hspre\(\let\hspost\)%
   \pboxed}%
  {\endpboxed%
   \hsnewpar\belowdisplayskip
   \ignorespacesafterend}

\newenvironment{oldplainhscode}%
  {\hsnewpar\abovedisplayskip
   \advance\leftskip\mathindent
   \hscodestyle
   \let\\=\@normalcr
   \(\pboxed}%
  {\endpboxed\)%
   \hsnewpar\belowdisplayskip
   \ignorespacesafterend}

% Here, we make plainhscode the default environment.

\newcommand{\plainhs}{\sethscode{plainhscode}}
\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
\plainhs

% The arrayhscode is like plain, but makes use of polytable's
% parray environment which disallows page breaks in code blocks.

\newenvironment{arrayhscode}%
  {\hsnewpar\abovedisplayskip
   \advance\leftskip\mathindent
   \hscodestyle
   \let\\=\@normalcr
   \(\parray}%
  {\endparray\)%
   \hsnewpar\belowdisplayskip
   \ignorespacesafterend}

\newcommand{\arrayhs}{\sethscode{arrayhscode}}

% The mathhscode environment also makes use of polytable's parray 
% environment. It is supposed to be used only inside math mode 
% (I used it to typeset the type rules in my thesis).

\newenvironment{mathhscode}%
  {\parray}{\endparray}

\newcommand{\mathhs}{\sethscode{mathhscode}}

% texths is similar to mathhs, but works in text mode.

\newenvironment{texthscode}%
  {\(\parray}{\endparray\)}

\newcommand{\texths}{\sethscode{texthscode}}

% The framed environment places code in a framed box.

\def\codeframewidth{\arrayrulewidth}
\RequirePackage{calc}

\newenvironment{framedhscode}%
  {\parskip=\abovedisplayskip\par\noindent
   \hscodestyle
   \arrayrulewidth=\codeframewidth
   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
   \hline\framedhslinecorrect\\{-1.5ex}%
   \let\endoflinesave=\\
   \let\\=\@normalcr
   \(\pboxed}%
  {\endpboxed\)%
   \framedhslinecorrect\endoflinesave{.5ex}\hline
   \endtabular
   \parskip=\belowdisplayskip\par\noindent
   \ignorespacesafterend}

\newcommand{\framedhslinecorrect}[2]%
  {#1[#2]}

\newcommand{\framedhs}{\sethscode{framedhscode}}

% The inlinehscode environment is an experimental environment
% that can be used to typeset displayed code inline.

\newenvironment{inlinehscode}%
  {\(\def\column##1##2{}%
   \let\>\undefined\let\<\undefined\let\\\undefined
   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
   \def\fromto##1##2##3{##3}%
   \def\nextline{}}{\) }%

\newcommand{\inlinehs}{\sethscode{inlinehscode}}

% The joincode environment is a separate environment that
% can be used to surround and thereby connect multiple code
% blocks.

\newenvironment{joincode}%
  {\let\orighscode=\hscode
   \let\origendhscode=\endhscode
   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
   %\let\SaveRestoreHook=\empty
   %\let\ColumnHook=\empty
   %\let\resethooks=\empty
   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
  {\origendhscode
   \global\let\hscode=\orighscode
   \global\let\endhscode=\origendhscode}%

\makeatother
\EndFmtInput
%

@@

  Suppose that we would like to update a list at a specific position:
\begin{hscode}\SaveRestoreHook
~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Int}~\WithMath{->}{\rightarrow}~[a]~\WithMath{->}{\rightarrow}~[a]
\\
\texfamily ~~modify~f~i~xs~=~helper~[]~xs~0
\\
\texfamily ~~~{\bfseries where}~helper~before~(x~:~after)~!j
\\
\texfamily ~~~~~~~~~~~|~j~\WithMath{==}{\equiv}~i~~~~=~before~\WithMath{++}{+\!\!+}~f~x~\WithMath{++}{+\!\!+}~after
\\
\texfamily ~~~~~~~~~~~|~\textit{otherwise}~=~helper~(before~\WithMath{++}{+\!\!+}~[x])~after~(j~+~1)
\\
\texfamily ~~~~~~~~~helper~\char95 ~[]~\char95 ~=~\textit{error}~\char34 Index~out~of~bounds.\char34 
\ColumnHook
\end{hscode}\resethooks
  Here \text{\tt modify} takes an update action \text{\tt f}\footnote{\text{\tt f} returns a list rather
  than a single element to prevent curious readers from suggesting to use a
  boxed array instead of a list.}, an index \text{\tt i}, and a list \text{\tt xs} and returns a
  new list with the \text{\tt i}'th element replaced with the result of \text{\tt f}.%
  This function "unpacks" a list, modifies one element, and "packs" the result
  into a list. If we do a lot of updates, we end up unpacking and packing the
  list over and over again -- very time-consuming for long lists. Explicitly
  working with the unpacked representation is bug-prone. A list zipper
  simplifies this.

  A zipper consists of a focus (alternatively called a hole) and surrounding
  context:
\begin{hscode}\SaveRestoreHook
~~{\bfseries data}~{\itshape Zipper}~a~=~{\itshape Zipper}~\char123 ~\char95 hole~::~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\\
\texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Context}~[a]~[a]
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\ColumnHook
\end{hscode}\resethooks
  where the \text{\tt ListContext} keeps track of elements to the left and to the right
  of the focus. We can now define movements:
\begin{hscode}\SaveRestoreHook
~~left~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~left~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~[]~\char95 ))~~~~~~~~=~{\itshape Nothing}
\\
\texfamily ~~left~({\itshape Zipper}~hole~({\itshape Context}~(l~:~ls)~rs))~=~{\itshape Just}~\$~{\itshape Zipper}~l~\$~{\itshape Context}~ls~(hole~:~rs)
\\
\texfamily 
\\
\texfamily ~~right~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~right~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~\char95 ~[]))~~~~~~~~=~{\itshape Nothing}
\\
\texfamily ~~right~({\itshape Zipper}~hole~({\itshape Context}~ls~(r~:~rs)))~=~{\itshape Just}~\$~{\itshape Zipper}~r~\$~{\itshape Context}~(hole~:~ls)~rs
\ColumnHook
\end{hscode}\resethooks
  and functions for entering and leaving the zipper:
\begin{hscode}\SaveRestoreHook
~~lzEnter~::~[a]~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~lzEnter~[]~~~~~~~=~{\itshape Nothing}
\\
\texfamily ~~lzEnter~(x~:~xs)~=~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~[]~xs)
\\
\texfamily 
\\
\texfamily ~~lzLeave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~[a]
\\
\texfamily ~~lzLeave~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~reverse~ls~\WithMath{++}{+\!\!+}~hole~:~rs
\ColumnHook
\end{hscode}\resethooks

  Finally, we define a local version of our \text{\tt modify} function (\TODO{Boy, is
  this function ugly...})
\begin{hscode}\SaveRestoreHook
~~lzModify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~lzModify~f~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~{\bfseries case}~f~hole~{\bfseries of}
\\
\texfamily ~~~~(x~:~xs)~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~ls~(xs~\WithMath{++}{+\!\!+}~rs))
\\
\texfamily ~~~~[]~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~rs~{\bfseries of}
\\
\texfamily ~~~~~~(r~:~rs')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape Context}~ls~rs')
\\
\texfamily ~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~ls~{\bfseries of}
\\
\texfamily ~~~~~~~~(l~:~ls')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~l~({\itshape Context}~ls'~rs)
\\
\texfamily ~~~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Nothing}
\ColumnHook
\end{hscode}\resethooks

  \newpage
  using which we can perform multiple updates efficiently and with minimal code
  bloat\footnote{%
  Operator \text{\tt \char62{}\char61{}\char62{}} comes from \text{\tt Control\char46{}Monad} module in \text{\tt base}
  and has the following signature:
\begin{hscode}\SaveRestoreHook
~~(>=>)~::~{\itshape Monad}~m~=>~(a~\WithMath{->}{\rightarrow}~m~b)~\WithMath{->}{\rightarrow}~(b~\WithMath{->}{\rightarrow}~m~c)~\WithMath{->}{\rightarrow}~a~\WithMath{->}{\rightarrow}~m~c
\ColumnHook
\end{hscode}\resethooks
  }:
\begin{hscode}\SaveRestoreHook
~~modifyExample~::~{\itshape IO}~()
\\
\texfamily ~~modifyExample~=~print~\$
\\
\texfamily ~~~~lzEnter~@{\itshape Int}~>=>~right
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~right
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(const~[])
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(return~\WithMath{.}{\circ}~(+1))
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~left
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(return~\WithMath{.}{\circ}~negate)
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~return~\WithMath{.}{\circ}~lzLeave~\$
\\
\texfamily ~~~~~~[1,~2,~3,~4,~5]
\ColumnHook
\end{hscode}\resethooks

  Consider now a binary tree data structure:
\begin{hscode}\SaveRestoreHook
~~{\bfseries data}~{\itshape Tree}~a
\\
\texfamily ~~~~=~{\itshape Fork}~({\itshape Tree}~a)~({\itshape Tree}~a)
\\
\texfamily ~~~~|~{\itshape Leaf}~!a
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\ColumnHook
\end{hscode}\resethooks
  A binary tree zipper is slightly more insteresting than the list zipper,
  because we can move up and down the tree as well. The zipper again consists of
  a hole (a subtree we are focused on) and its surrounding context (a path from
  the hole to the root of the tree):
\begin{hscode}\SaveRestoreHook
~~{\bfseries data}~{\itshape Zipper}~a~=~{\itshape Zipper}~\char123 ~\char95 hole~::~({\itshape Tree}~a),~\char95 cxt~::~!({\itshape Context}~a)~\char125 
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\\
\texfamily 
\\
\texfamily ~~{\bfseries data}~{\itshape Context}~a
\\
\texfamily ~~~~=~{\itshape Top}
\\
\texfamily ~~~~|~{\itshape Left}~!({\itshape Context}~a)~({\itshape Tree}~a)
\\
\texfamily ~~~~|~{\itshape TreeRight}~({\itshape Tree}~a)~!({\itshape Context}~a)
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\ColumnHook
\end{hscode}\resethooks
  To move the zipper down, we "unpack" the current hole:
\begin{hscode}\SaveRestoreHook
~~down~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~down~({\itshape Zipper}~({\itshape Leaf}~\char95 )~\char95 )~=~{\itshape Nothing}
\\
\texfamily ~~down~({\itshape Zipper}~({\itshape Fork}~l~r)~cxt)~=~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape TreeRight}~l~cxt)
\ColumnHook
\end{hscode}\resethooks
  \text{\tt TreeContext} stores everything we need to reconstruct the hole, and \text{\tt tzUp} does
  exactly that:
\begin{hscode}\SaveRestoreHook
~~up~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~up~({\itshape Zipper}~\char95 ~{\itshape Top})~=~{\itshape Nothing}
\\
\texfamily ~~up~({\itshape Zipper}~l~({\itshape Left}~cxt~r))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
\\
\texfamily ~~up~({\itshape Zipper}~r~({\itshape TreeRight}~l~cxt))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
\ColumnHook
\end{hscode}\resethooks
  Implementation of \text{\tt tzLeft}, \text{\tt tzRight}, and \text{\tt tzEnter} is very similar to the
  List zipper case and is left as an exercise for the reader. \text{\tt tzLeave} differs
  slightly in that we now move all the way up rather than left:
\begin{hscode}\SaveRestoreHook
~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Tree}~a
\\
\texfamily ~~leave~z~=~{\bfseries case}~up~z~{\bfseries of}
\\
\texfamily ~~~~{\itshape Just}~z'~\WithMath{->}{\rightarrow}~leave~z'
\\
\texfamily ~~~~{\itshape Nothing}~\WithMath{->}{\rightarrow}~\char95 hole~z
\ColumnHook
\end{hscode}\resethooks

@@

  The generic zipper we will use is very similar to the one presented
  in\cite{adams2010syz}. The most common technique in Haskell for supporting
  heterogeneous types is Existential Quantification. However, not every type can
  act as a hole. To support moving down the tree, we need the hole to be
  \textit{dissectible}, i.e. we would like to be able to dissect the value into
  the constructor and its arguments. Even though \text{\tt Data\char46{}Data\char46{}gfold} allows us to achieve
  this, we define out own typeclass which additionally allows us to propagate
  down arbitrary constraints:
\begin{hscode}\SaveRestoreHook
~~{\bfseries class}~{\itshape Dissectible}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(a~::~{\itshape Type})~{\bfseries where}
\\
\texfamily ~~~~dissect~::~a~\WithMath{->}{\rightarrow}~{\itshape Left}~c~a
\\
\texfamily 
\\
\texfamily ~~{\bfseries data}~{\itshape Left}~c~expects~{\bfseries where}
\\
\texfamily ~~~~{\itshape LOne}~~::~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~b
\\
\texfamily ~~~~{\itshape LCons}~::~(c~b,~{\itshape Dissectible}~c~b)~=>~{\itshape Left}~c~(b~\WithMath{->}{\rightarrow}~expects)~\WithMath{->}{\rightarrow}~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~expects
\ColumnHook
\end{hscode}\resethooks
  For example, here is how we can make \text{\tt Tree} an instance of \text{\tt Disssectible}
\begin{hscode}\SaveRestoreHook
~~{\bfseries instance}~c~({\itshape Tree}~a)~=>~{\itshape Dissectible}~c~({\itshape Tree}~a)~{\bfseries where}
\\
\texfamily ~~~~dissect~({\itshape Fork}~l~r)~=~{\itshape LOne}~{\itshape Fork}~`{\itshape LCons}`~l~`{\itshape LCons}`~r
\\
\texfamily ~~~~dissect~x~=~{\itshape LOne}~x
\ColumnHook
\end{hscode}\resethooks
  We can unpack a \text{\tt Fork} and the zipper will thus be able to go down. \text{\tt Leaf}s,
  however, are left untouched and trying to go down from a \text{\tt Leaf} will return
  \text{\tt Nothing}.

>>>>

\EndFmtInput
\makeatother
%
%

@@

  Suppose that we would like to update a list at a specific position:
\begin{tabbing}\texfamily
~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Int}~\WithMath{->}{\rightarrow}~[a]~\WithMath{->}{\rightarrow}~[a]
\\
\texfamily ~~modify~f~i~xs~=~helper~[]~xs~0
\\
\texfamily ~~~{\bfseries where}~helper~before~(x~:~after)~!j
\\
\texfamily ~~~~~~~~~~~|~j~\WithMath{==}{\equiv}~i~~~~=~before~\WithMath{++}{+\!\!+}~f~x~\WithMath{++}{+\!\!+}~after
\\
\texfamily ~~~~~~~~~~~|~\textit{otherwise}~=~helper~(before~\WithMath{++}{+\!\!+}~[x])~after~(j~+~1)
\\
\texfamily ~~~~~~~~~helper~\char95 ~[]~\char95 ~=~\textit{error}~\char34 Index~out~of~bounds.\char34 
\end{tabbing}
  Here \text{\ttfamily modify} takes an update action \text{\ttfamily f}\footnote{\text{\ttfamily f} returns a list rather
  than a single element to prevent curious readers from suggesting to use a
  boxed array instead of a list.}, an index \text{\ttfamily i}, and a list \text{\ttfamily xs} and returns a
  new list with the \text{\ttfamily i}'th element replaced with the result of \text{\ttfamily f}.%
  This function "unpacks" a list, modifies one element, and "packs" the result
  into a list. If we do a lot of updates, we end up unpacking and packing the
  list over and over again -- very time-consuming for long lists. Explicitly
  working with the unpacked representation is bug-prone. A list zipper
  simplifies this.

  A zipper consists of a focus (alternatively called a hole) and surrounding
  context:
\begin{tabbing}\texfamily
~~{\bfseries data}~{\itshape Zipper}~a~=~{\itshape Zipper}~\char123 ~\char95 hole~::~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\\
\texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Context}~[a]~[a]
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\end{tabbing}
  where the \text{\ttfamily ListContext} keeps track of elements to the left and to the right
  of the focus. We can now define movements:
\begin{tabbing}\texfamily
~~left~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~left~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~[]~\char95 ))~~~~~~~~=~{\itshape Nothing}
\\
\texfamily ~~left~({\itshape Zipper}~hole~({\itshape Context}~(l~:~ls)~rs))~=~{\itshape Just}~\$~{\itshape Zipper}~l~\$~{\itshape Context}~ls~(hole~:~rs)
\\
\texfamily 
\\
\texfamily ~~right~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~right~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~\char95 ~[]))~~~~~~~~=~{\itshape Nothing}
\\
\texfamily ~~right~({\itshape Zipper}~hole~({\itshape Context}~ls~(r~:~rs)))~=~{\itshape Just}~\$~{\itshape Zipper}~r~\$~{\itshape Context}~(hole~:~ls)~rs
\end{tabbing}
  and functions for entering and leaving the zipper:
\begin{tabbing}\texfamily
~~lzEnter~::~[a]~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~lzEnter~[]~~~~~~~=~{\itshape Nothing}
\\
\texfamily ~~lzEnter~(x~:~xs)~=~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~[]~xs)
\\
\texfamily 
\\
\texfamily ~~lzLeave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~[a]
\\
\texfamily ~~lzLeave~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~reverse~ls~\WithMath{++}{+\!\!+}~hole~:~rs
\end{tabbing}

  Finally, we define a local version of our \text{\ttfamily modify} function (\TODO{Boy, is
  this function ugly...})
\begin{tabbing}\texfamily
~~lzModify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~lzModify~f~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~{\bfseries case}~f~hole~{\bfseries of}
\\
\texfamily ~~~~(x~:~xs)~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~ls~(xs~\WithMath{++}{+\!\!+}~rs))
\\
\texfamily ~~~~[]~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~rs~{\bfseries of}
\\
\texfamily ~~~~~~(r~:~rs')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape Context}~ls~rs')
\\
\texfamily ~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~ls~{\bfseries of}
\\
\texfamily ~~~~~~~~(l~:~ls')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~l~({\itshape Context}~ls'~rs)
\\
\texfamily ~~~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Nothing}
\end{tabbing}

  \newpage
  using which we can perform multiple updates efficiently and with minimal code
  bloat\footnote{%
  Operator \text{\ttfamily \char62{}\char61{}\char62{}} comes from \text{\ttfamily Control\char46{}Monad} module in \text{\ttfamily base}
  and has the following signature:
\begin{tabbing}\texfamily
~~(>=>)~::~{\itshape Monad}~m~=>~(a~\WithMath{->}{\rightarrow}~m~b)~\WithMath{->}{\rightarrow}~(b~\WithMath{->}{\rightarrow}~m~c)~\WithMath{->}{\rightarrow}~a~\WithMath{->}{\rightarrow}~m~c
\end{tabbing}
  }:
\begin{tabbing}\texfamily
~~modifyExample~::~{\itshape IO}~()
\\
\texfamily ~~modifyExample~=~print~\$
\\
\texfamily ~~~~lzEnter~@{\itshape Int}~>=>~right
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~right
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(const~[])
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(return~\WithMath{.}{\circ}~(+1))
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~left
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(return~\WithMath{.}{\circ}~negate)
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~return~\WithMath{.}{\circ}~lzLeave~\$
\\
\texfamily ~~~~~~[1,~2,~3,~4,~5]
\end{tabbing}

  Consider now a binary tree data structure:
\begin{tabbing}\texfamily
~~{\bfseries data}~{\itshape Tree}~a
\\
\texfamily ~~~~=~{\itshape Fork}~({\itshape Tree}~a)~({\itshape Tree}~a)
\\
\texfamily ~~~~|~{\itshape Leaf}~!a
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\end{tabbing}
  A binary tree zipper is slightly more insteresting than the list zipper,
  because we can move up and down the tree as well. The zipper again consists of
  a hole (a subtree we are focused on) and its surrounding context (a path from
  the hole to the root of the tree):
\begin{tabbing}\texfamily
~~{\bfseries data}~{\itshape Zipper}~a~=~{\itshape Zipper}~\char123 ~\char95 hole~::~({\itshape Tree}~a),~\char95 cxt~::~!({\itshape Context}~a)~\char125 
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\\
\texfamily 
\\
\texfamily ~~{\bfseries data}~{\itshape Context}~a
\\
\texfamily ~~~~=~{\itshape Top}
\\
\texfamily ~~~~|~{\itshape Left}~!({\itshape Context}~a)~({\itshape Tree}~a)
\\
\texfamily ~~~~|~{\itshape TreeRight}~({\itshape Tree}~a)~!({\itshape Context}~a)
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\end{tabbing}
  To move the zipper down, we "unpack" the current hole:
\begin{tabbing}\texfamily
~~down~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~down~({\itshape Zipper}~({\itshape Leaf}~\char95 )~\char95 )~=~{\itshape Nothing}
\\
\texfamily ~~down~({\itshape Zipper}~({\itshape Fork}~l~r)~cxt)~=~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape TreeRight}~l~cxt)
\end{tabbing}
  \text{\ttfamily TreeContext} stores everything we need to reconstruct the hole, and \text{\ttfamily tzUp} does
  exactly that:
\begin{tabbing}\texfamily
~~up~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~up~({\itshape Zipper}~\char95 ~{\itshape Top})~=~{\itshape Nothing}
\\
\texfamily ~~up~({\itshape Zipper}~l~({\itshape Left}~cxt~r))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
\\
\texfamily ~~up~({\itshape Zipper}~r~({\itshape TreeRight}~l~cxt))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
\end{tabbing}
  Implementation of \text{\ttfamily tzLeft}, \text{\ttfamily tzRight}, and \text{\ttfamily tzEnter} is very similar to the
  List zipper case and is left as an exercise for the reader. \text{\ttfamily tzLeave} differs
  slightly in that we now move all the way up rather than left:
\begin{tabbing}\texfamily
~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Tree}~a
\\
\texfamily ~~leave~z~=~{\bfseries case}~up~z~{\bfseries of}
\\
\texfamily ~~~~{\itshape Just}~z'~\WithMath{->}{\rightarrow}~leave~z'
\\
\texfamily ~~~~{\itshape Nothing}~\WithMath{->}{\rightarrow}~\char95 hole~z
\end{tabbing}

@@

  The generic zipper we will use is very similar to the one presented
  in\cite{adams2010syz}. The most common technique in Haskell for supporting
  heterogeneous types is Existential Quantification. However, not every type can
  act as a hole. To support moving down the tree, we need the hole to be
  \textit{dissectible}, i.e. we would like to be able to dissect the value into
  the constructor and its arguments. Even though \text{\ttfamily Data\char46{}Data\char46{}gfold} allows us to achieve
  this, we define out own typeclass which additionally allows us to propagate
  down arbitrary constraints:
\begin{tabbing}\texfamily
~~{\bfseries class}~{\itshape Dissectible}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(a~::~{\itshape Type})~{\bfseries where}
\\
\texfamily ~~~~dissect~::~a~\WithMath{->}{\rightarrow}~{\itshape Left}~c~a
\\
\texfamily 
\\
\texfamily ~~{\bfseries data}~{\itshape Left}~c~expects~{\bfseries where}
\\
\texfamily ~~~~{\itshape LOne}~~::~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~b
\\
\texfamily ~~~~{\itshape LCons}~::~(c~b,~{\itshape Dissectible}~c~b)~=>~{\itshape Left}~c~(b~\WithMath{->}{\rightarrow}~expects)~\WithMath{->}{\rightarrow}~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~expects
\end{tabbing}
  For example, here is how we can make \text{\ttfamily Tree} an instance of \text{\ttfamily Disssectible}
\begin{tabbing}\texfamily
~~{\bfseries instance}~c~({\itshape Tree}~a)~=>~{\itshape Dissectible}~c~({\itshape Tree}~a)~{\bfseries where}
\\
\texfamily ~~~~dissect~({\itshape Fork}~l~r)~=~{\itshape LOne}~{\itshape Fork}~`{\itshape LCons}`~l~`{\itshape LCons}`~r
\\
\texfamily ~~~~dissect~x~=~{\itshape LOne}~x
\end{tabbing}
  We can unpack a \text{\ttfamily Fork} and the zipper will thus be able to go down. \text{\ttfamily Leaf}s,
  however, are left untouched and trying to go down from a \text{\ttfamily Leaf} will return
  \text{\ttfamily Nothing}.
